---
title: "Assignment_9"
author: "mitchell"
date: "`r Sys.Date()`"
output: html_document
---

```{r, echo=FALSE, include=FALSE}

library(tidyverse)
library(tidyr)
library(readxl)
library(janitor)
library(lubridate)
library(gganimate)
library(transformr)
library(ggplot2)
library(stringr)
library(gganimate)
library(gifski)
library(skimr)
```
```{r start, echo=FALSE}
GradData <- read.csv("../../Data/GradSchool_Admissions.csv")

```

## R Markdown_GradSchool_Admissions
The following data analysis examines the acceptance rates of graduate school applicants. This analysis will help students visualize and predict which rank of school they are likely to be accepted to based on their GRE scores and GPA. I have used R markdown to present my data analysis in an HTML format.
 

```{r GradSchool_data, echo=FALSE}

skim(GradData)
# summary(GradData)
```
The source data has 4 variables and 400 observations. The four variables include Admitance,gre,gpa,rank.

## The story told in Plots
```{r pressure, echo=FALSE}
heatmap <- ggplot(GradData, aes(x = gre, y = gpa, fill = admit)) +
  geom_tile() +
  scale_fill_gradient(low = "red", high = "green") +
  labs(title = "Heatmap of GRE vs. GPA by Acceptance",
       x = "GRE Score",
       y = "GPA",
       fill = "Acceptance") +
  theme_minimal()

# Print the heatmap
print(heatmap)

```

## Bar chart for "admitance" which is a binary outcome
```{r binary outcome predictiors, echo=FALSE}
barplot(table(GradData$admit), col = c("red", "green"),
        xlab = "Admissions Status", names.arg = c("Not Admitted", "Admitted"))

```

## Histograms for GRE Scores and GPA, which are numerical predictors
```{r numerical predictors, echo=FALSE}

hist(GradData$gre, col = "lightpink", xlab = "GRE Score", main = "Histogram of GRE Scores")
hist(GradData$gpa, col = "purple", xlab = "GPA", main = "Histogram of GPAs")


```

## Box plot to visualize the relationship between GRE scores and admission
```{r boxplot, echo=FALSE}
# Box plot to visualize the relationship between GRE scores and admission
plot1 <- ggplot(GradData, aes(x = admit, y = gre, fill = factor(admit))) +
  geom_boxplot() +
  labs(title = "GRE Scores vs. Admission",
       x = "Admission Status",
       y = "GRE Score") +
  scale_fill_manual(values = c("red", "green")) +
  theme(legend.key.size = unit(1.5, "cm"))

print(plot1)
  

# Box plot to visualize the relationship between GPA scores and admission
plot2 <- boxplot(GradData$gpa ~ GradData$admit, col = c("red", "green"),
        names = c("Not Admitted", "Admitted"),
        main = "GPA Scores vs. Admission",
        ylab = "GPA Score",
        xlab = "Admission Status")

```
  
## Scatter Plot by School Rank  
A scatter plot of GRE scores versus GPAs was created for all rankings.
 
```{r scatter_plot_ALL, echo=FALSE}

scatter_plot <- ggplot(GradData, aes(x = gre, y = gpa, color = factor(rank))) +
  geom_point() +
  labs(title = "Scatter Plot of GRE vs. GPA by Rank",
       x = "GRE Score",
       y = "GPA",
       color = "Rank") +
  scale_color_manual(values = c("black", "black", "black", "black")) +
  theme_minimal()
print(scatter_plot)
```

Rank 3 was chosen for further investigation of the likelihood of a linear relationship between GPA and GRE score using a linear regression model.

```{r scatter_plot_rank3, echo=FALSE}
# rank_filter
selected_rank <- 3

# Scatter plot filtered by rank
scatter_plot <- ggplot(GradData, aes(x = gre, y = gpa, color = factor(rank))) +
  geom_point(data = subset(GradData, rank == selected_rank)) +  # Filter the data
  labs(title = paste("Scatter Plot of GRE vs. GPA for Rank", selected_rank),
       x = "GRE Score",
       y = "GPA",
       color = "Rank") +
  scale_color_manual(values = c("green")) +
  theme_minimal()

# Print the scatter plot
print(scatter_plot)
```

## MODEL 
The value of the actual data observed for school ranked # 3 can be compared to the model of the scatter plot by plotting the point (3, value of the actual data observed for school ranked # 3) on the scatter plot.

```{r, mitch, echo=FALSE, warning=FALSE}
scatter_plot_with_model <- scatter_plot +
  geom_smooth(method = "glm", formula = y ~ x, data = GradData,
              method.args = list(family = binomial), color = "limegreen", linetype = "dashed")

scatter_plot_with_model


```


The model shows the predictors of graduate school admission 


GRE vs. Admit

```{r gre/admit , echo=FALSE}

plot(GradData$gre, GradData$admit, main = "GRE vs. Admission Status",
     xlab = "GRE Score", ylab = "Admission Status")

```



GPA vs. Admit

```{r, GPA vs. Admit, echo=FALSE}

plot(GradData$gpa, GradData$admit, main = "GPA vs. Admission Status",
     xlab = "GPA", ylab = "Admission Status")
```
```

```{r model, echo=FALSE}
model1 <- glm(admit ~ gpa, data = GradData, family = binomial)
model2 <- lm(gre ~ gpa, data = GradData)

```

```{r summary, echo=TRUE}
summary(model1)
summary(model2)
```


### Conclusion:
The data analysis shows that there is a strong correlation between GRE scores and GPA, and acceptance rates to graduate schools. Students with higher GRE scores and GPAs are more likely to be accepted to higher-ranked schools
v n
In conclusion, the following exploratory plots and the logistic regression model show that GRE scores and GPA appear to be positively associated with admission.

A higher rank (lower number) of the undergraduate institution indicates a higher probability of admission.

The logistic regression model provides insights into the significance of these variables and how they collectively affect the likelihood of admission.


