---
title: "MILLERBERG_Skills_Test_3"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(skimr)
library(janitor)
library(tidyr)
library(broom)
library(tidyverse)
library(readxl)
library(lubridate)
library(gganimate)
library(transformr)
library(stringr)
library(gganimate)
library(gifski)
library(dplyr)
library(sf)
library(gplots)
library(scales)
library(easystats)
library(ggplot2)
```

# Skills test
This report aims to demonstrate proficiency in data manipulation, statistical modeling, and visualization using R. The tasks involve working with two datasets, "FacultySalaries_1995.csv" and "Juniper_Oils.csv," with an emphasis on data cleaning, exploratory data analysis, and statistical modeling.

# Task 1: "FacultySalaries_1995.csv"
The first task involves loading and cleaning the "FacultySalaries_1995.csv" dataset. This dataset contains information about college faculty salaries, total compensation, ranks, tiers, and states. The objective is to create a tidy dataset and reproduce a specified graph. College faculty ranks include Assistant, Associate, and Full, while tiers represent the funding allocation for research and teaching.


# Reading Data:

The code creates a file path and reads in a CSV file named "FacultySalaries_1995.csv" into a data frame named data1. The read.csv function is used to read the CSV file.
```{r}

file_path <- "FacultySalaries_1995.csv"

data1 <- read.csv(file_path)


```


# Cleaning Data:

Using the pivot_longer function from the tidyverse package to transform the data from wide to long format. The next step in cleaning this data was to create the salary-related columns (AvgFullProfSalary, AvgAssocProfSalary, AvgAssistProfSalary). These columns have been combined into a single column named "Salary" with a corresponding "Rank" column indicating the type of faculty. Similar transformations are applied to compensation and number-related columns.

```{r}

data1 <- data1 %>% 
  pivot_longer(cols = c("AvgFullProfSalary", "AvgAssocProfSalary", "AvgAssistProfSalary"),
               names_to = "Rank",
               values_to = "Salary") %>% 
  pivot_longer(cols = c("AvgFullProfComp", "AvgAssocProfComp", "AvgAssistProfComp"),
               names_to = "Rank_Comp",
               values_to = "Compensation") %>% 
  pivot_longer(cols = c("NumFullProfs", "NumAssocProfs", "NumAssistProfs", "NumInstructors"),
               names_to = "Rank_Num",
               values_to = "Staff_Number") %>% 
  mutate(Rank = case_when(Rank == "AvgFullProfSalary" ~ "Full",
                          Rank == "AvgAssocProfSalary" ~ "Assoc",
                          Rank == "AvgAssistProfSalary" ~ "Assist"))


```

# Filtering Data:

Removes rows where the "Tier" variable is equal to "VIIB".

```{r}

data1 <- data1[!(data1$Tier=="VIIB"), ]
```

# Visualization Data:

This code uses ggplot2 to create a boxplot of salary distribution (Salary) across different faculty ranks (Rank), with each rank having a different fill color.
The facet_wrap function is used to create separate boxplots for each tier (Tier), arranging them in a grid.
Various theme functions are applied to modify the appearance of the plot, such as adjusting legend size, text size, axis labels, and strip text.

```{r pt.2, warning=FALSE}

data1 %>% 
  ggplot(aes(x=Rank,y=Salary)) +
  geom_boxplot(aes(fill=Rank)) +
  facet_wrap(~Tier) +
  theme_minimal() +
  theme(legend.key.size = unit(1, "cm"),
        legend.title = element_text(size = 19, color = "black"),
        legend.text = element_text(size = 13, color = "black"),
        axis.text.x = element_text(angle = 65, size = 13, color = "black"),
        axis.text.y = element_text(size = 13, color = "black"),
        axis.title.x = element_text(size = 19, color = "black"),
        axis.title.y = element_text(size = 19, color = "black"),
        strip.text.x = element_text(size = 13, color = "black"))

```

# Exporting Figure:

The ggsave function is then used to save the created plot as a JPEG image named "millerberg_Fig_1.jpg".

```{r warning=FALSE}
ggsave("./millerberg_Fig_1.jpg")
```

# Task 2: ANOVA Modeling
In the second task, an analysis of variance (ANOVA) model is constructed to examine the influence of "State," "Tier," and "Rank" on "Salary." The ANOVA model tests the main effects of these predictors without including interactions.

# Building and Analyzing ANOVA Model:

This code creates a generalized linear model (GLM) where Salary is modeled as a function  of Rank, State, and Tier in the dataset data1. The summary function of the fitted GLM model, shows coefficients, standard errors, t-values, and p-values.

```{r}
model1 <- glm(data = data1, 
            formula = Salary ~ Rank + State + Tier )
summary(model1)
```

# Exporting ANOVA Summary:

The sink function diverts the output to a file ("Salary_ANOVA_Summary.txt") instead of the console, providing a written summary of the ANOVA results.

```{r pt.3,}

anova(model1)
summary(model1)

sink("./Salary_ANOVA_Summary.txt") 
anova(model1)
sink(NULL)

```


# Task 3: "Juniper_Oils.csv"

Moving to the "Juniper_Oils.csv" dataset, the third task involves loading and tidying the data. This dataset contains measurements of chemical concentrations in the essential oil of cedar trees over various years since a burn event.

# Cleaning and Tidying Another Dataset:
Loading and Tidying Data:
This code reads the data from "Juniper_Oils.csv" and tidies it by converting chemical concentrations from wide to long format using pivot_longer.

```{r pt.4,}
data2 <- read_csv("./Juniper_Oils.csv")

data2 <- data2 %>% 
  pivot_longer(cols = c("alpha-pinene","para-cymene","alpha-terpineol","cedr-9-ene",
                        "alpha-cedrene","beta-cedrene","cis-thujopsene","alpha-himachalene",
                        "beta-chamigrene","cuparene","compound 1","alpha-chamigrene",
                        "widdrol","cedrol","beta-acorenol","alpha-acorenol","gamma-eudesmol",
                        "beta-eudesmol","alpha-eudesmol","cedr-8-en-13-ol","cedr-8-en-15-ol",
                        "compound 2","thujopsenal"),
               names_to = "ChemicalID",
               values_to = "Concentration")
```

# Task 4: Graphical Representation
The fourth task requires creating a graph that visualizes the relationship between "YearsSinceBurn" and "Concentration" for different chemical compounds. 

# Data Preparation:
The code is using the %>% (pipe) operator to pass the data2 dataframe to the subsequent operations. The function of ggplot is to initializes the plotting object.
aes sets the aesthetic mappings, specifying that YearsSinceBurn should be on the x-axis and Concentration on the y-axis.
Facet_wrap creates a small multiple plot, faceting the data by ChemicalID. Each facet represents a different chemical.
scales = "free_y" allows the y-axes to have different scales for each facet.
Geom_smooth adds a smooth line to the plot using the loess method, which is a non-parametric regression.Theme_minimal() sets a minimal theme for the plot.Labs is used to set the x-axis label and the title.Theme(plot.title = element_text(hjust = 0.5)) centers the plot title.
```{r}
data2 %>% 
  ggplot(aes(x=YearsSinceBurn, y=Concentration)) +
  facet_wrap(~ChemicalID, scales = "free_y") +
  geom_smooth(method = loess) +
  theme_minimal() +
  labs(x = "Years Since Burn",
       title = "Concentration by Chemical") +
  theme(plot.title = element_text(hjust = 0.5))


```

# Exporting the Plot:

ggsave saves the plot as an image file named "millerberg_Fig_2.jpg" in the current working directory.
```{r}
# export to millerberg_Fig_2.jpg
ggsave("./millerberg_Fig_2.jpg")
```

## Task 5: Generalized Linear Model (GLM)
The final task entails applying a generalized linear model (GLM) to identify chemicals whose concentrations significantly change with "Years Since Burn." The focus is on chemicals with a p-value < 0.05. The results are presented in a tidy format using the tidy() function from the broom package.

# Generalized Linear Model (GLM):
This R code performs statistical analysis using a generalized linear model (GLM) to identify chemicals whose concentrations are significantly affected by "Years Since Burn." It then produces a tidy data frame containing information about statistically significant effects. glm is used to create a generalized linear model.
The formula Concentration ~ ChemicalID * YearsSinceBurn specifies that the concentration is modeled as a function of both the chemical ID and the interaction between chemical ID and years since burn. Summary extracts various pieces of information from the GLM, including coefficient estimates, standard errors, statistics, and p-values.

```{r}
model2 <- glm(data = data2, formula = Concentration ~ ChemicalID * YearsSinceBurn)

model2_sum <- summary(model2)

```

# Tidy Data Frame:
Tidy function from the broom package is used to tidy up the model results. It extracts and organizes information about the model parameters.

```{r}
tidy <- tidy(model2)

```

# Subset Tidy Data Frame:
This code creates a subset of the tidy data frame (tidy) containing only rows where the p-value is less than or equal to 0.05, indicating statistical significance. The print function then prints the resulting tidy data frame containing information about statistically significant effects.
```{r}
ttss <- tidy[c(tidy$p.value <= 0.05), ]
print(ttss)
```

